### 2018-05-03 4_impl.tex
Change font of representations names. Add mem-units descrition.

The general architecture scheme of \porthos[2] is presented in Figure~\ref{fig:arch}. On the picture, rectangles denote processing units (marked with gear sign with a number of the component for referencing it).

\begin{figure}%[H]
    \centering
  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/draw.io/general_arch.png}
  \caption{The general architecture of \porthos[2]}
  \label{fig:arch}
\end{figure}

@@

\subsection{The internal representations}
\label{ch:impl:model}
%As it has already been mentioned, all internal representations constructed by \porthos[2] are \textit{immutable}.

@@

The first internal representation that \porthos[2] uses is the \ytree{}, which represents a rather high-level AST.
The Apendix~\ref{apx:trees} represents the file tree of main classes that constitute the \ytree{} hierarchy (as the inheritance tree might be obvious for the C-like AST, we confine ourselves to presenting the classes file tree only).

@@

According to the C standard, \textit{"any statement may be preceded by a prefix that declares an identifier as a label name"}.
The \ytree{} statements of follow this rule, however they these labels are symbolic, and they need to be resolved at the pre-compilation stage.
Apart from the set of statements listed before, we define the \texttt{YFunctionDefinition} and its inheritor a litmus-specific declaration \texttt{YProcessDefinition}
%\textbf{(TODO: rename YProcessStatement->YProcessDefinition)} %TODO <--
used in intra-procedural analysis mode.
The function definition contains the \texttt{YCompoundStatement}-typed body and the signature of type \texttt{YMethodSignature}, which is used in the function resolution during the compilation stage. %TODO: pre-compilation?
The other litmus-specific statements are \texttt{YPreludeDefinition}
%\textbf{(TODO: rename YPreludeStatement -> YPreludeDefinition)} %TODO <--
that carries the list of \texttt{YStatement}-typed initial writes, and \texttt{YPostludeDefinition} 
%\textbf{(TODO: rename YPostludeStatement -> YPostludeDefinition)} %TODO <--
that carries the \texttt{YExpression}-typed binary expression to be asserted by the litmus test.

@@

The \ytree{} is compiled into the low-level event-based program representation \xgraph{}.
The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:event}.
In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
%.. goto: compilation

atomic action?

%However, the \xgraph{}
%The basic element of \xgraph{} is an event. Each event \texttt{XEvent}
Firstly, two elements
%two basic elements of diff nature: XEvent and YMemoryUnit

comp event emulated by both -- optimisation?

@@

XEventInfo

hierarchy of memory units

@@

Valid graph:
- single source, two sinks
- connected
- each node has 1 or 2 children
- branching node (2 children) instanceof XComputationEvent (evaluation of the value, no shared-memory operations)
-

@@

@@

%\textit{Type analysis} is the process of \textit{converting} the text-based type labels (obtained at variable declarations) to the type instances (modelled by the class \texttt{mousquetaires.languages.common.Type}) and \textit{propagating} them to all expressions.
The C language has a static (resolved at compile-time) manifest (all types are declared explicitly) type system.
Comparing to languages that use type inference, the type analysis of a C program constitutes a simple propagating the type information (obtained from variables declarations) to all expressions.
In \porthos[2], the type is modelled by the class \texttt{mousquetaires.languages.common.Type}.%TODO: + fix class and package name!
Being carried at Y and X representation levels, the type is converted to a Z-type at the Z-formula encoding stage (see Section~\ref{ch:impl:proc:z-encoder}).

%TODO: describe the Type
%TODO: maybe rename back to XYType
%TODO: but definitely get rid of bit32

>>>>

The general architecture scheme of \porthos[2] is presented in Figure~\ref{fig:arch}. On the picture, rectangles denote processing units (marked with gear sign with a unique number of the component).

\begin{figure}%[H]
    \centering
  %draft=true
  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/draw.io/general_arch.png}
  \caption{The general architecture of \porthos[2]}
  \label{fig:arch}
\end{figure}

@@

\subsection{The internal representations}
\label{ch:impl:model}

As it has already been mentioned, all internal representations used by \porthos[2] are immutable.
For keeping the architecture transparent, we build all abstraction levels with interfaces, even if some of them does not add any new functionality.% (as java does not support multiple inheritance for classes (only for interfaces), we

@@

The first internal representation used by \porthos[2] is the \textit{\ytree{}}, which represents a rather high-level AST.
The Apendix~\ref{apx:trees} represents the file tree of main classes that constitute the \ytree{} hierarchy (as the inheritance tree might be obvious for the C-like AST, we confine ourselves to presenting the classes file tree only).

@@

According to the C standard, \textit{"any statement may be preceded by a prefix that declares an identifier as a label name"}.
The \ytree{} statements of follow this rule, however they these labels are symbolic, and they need to be resolved at the pre-compilation stage.
Apart from the set of statements listed before, we define the \texttt{YFunctionDefinition} and its inheritor a litmus-specific declaration \texttt{YProcessDefinition}
%\textbf{(TODO: rename YProcessStatement->YProcessDefinition)} %TODO <--
used in intra-procedural analysis mode.
%TODO: rename YMethodSignature -> YFunctionSignature
The function definition contains the \texttt{YCompoundStatement} body and the \texttt{YMethodSignature} signature, which is used in the function resolution during the compilation stage. %TODO: pre-compilation?
The other litmus-specific statements are \texttt{YPreludeDefinition}
%\textbf{(TODO: rename YPreludeStatement -> YPreludeDefinition)} %TODO <--
that carries the list of \texttt{YStatement} initial writes, and \texttt{YPostludeDefinition} 
%\textbf{(TODO: rename YPostludeStatement -> YPostludeDefinition)} %TODO <--
that carries the \texttt{YExpression} binary expression to be asserted by the litmus test.

@@

The \ytree{} is compiled into the low-level event-based program representation \textit{\xgraph{}}.
The mathematical structure of event-flow graph was discussed in Section~\ref{ch:wmm:event}.
%In general, \xgraph{} follows the structure described there, however it is constructed in several stages.
The nodes of the graph are events, and the edges are basic relations: the control-flow relation \po and the data-flow relations \co and \rf.
Hereinafter, we denote the \xgraph{} with only control-flow edges as \xgraph[CF], the \xgraph{} with only data-flow edges as \xgraph[CF]. The complete \xgraph{} is $\xgraph[CF+DF] = \xgraph[CF] \cup \xgraph[DF]$.
Figure~\ref{fig:class-diagrams:XEntity-interfaces} represents the main hierarchy of the X-abstraction level.

\begin{figure}[t]%[!b]%[H]
  \centering
  \includegraphics[width=\textwidth,keepaspectratio]{img/my/class-diagrams/XEntity-interfaces.png}
  \caption{The inheritance tree of main interfaces of \xgraph{} elements}
  \label{fig:class-diagrams:XEntity-interfaces}
\end{figure}

All elements of \xgraph{} implement the interface \texttt{XEntity}.
There are two main kinds of X-entity: \textit{events} that implement the \texttt{XEvent} interface, and \textit{memory-units} that implement the \texttt{XMemoryUnit} interface.

\paragraph{Memory units}
\label{ch:impl:model:xgraph:mem}

The \textit{memory unit} is a memory cell of an abstract machine executing the code.
This machine has an infinite number of arbitrary-sized \textit{registers} (local memory units) and \textit{locations} (shared memory units).
Local memory units inherit the \texttt{XProcessLocalElement} interface, that stores the ID of the owning process.
Figure~\ref{fig:class-diagrams:XMemoryUnit} represents inheritance hierarchy of memory units.

\begin{figure}[t]%[!b]%[H]
  \centering
  \includegraphics[width=\textwidth,height=\textheight,keepaspectratio]{img/my/class-diagrams/XMemoryUnit-m.png}
  \caption{The inheritance tree of \xgraph{} memory units}
  \label{fig:class-diagrams:XMemoryUnit}
\end{figure}

Following the terminology of the C standard, we distinguish the \textit{r-value} and \textit{l-value} memory units (l-values may be assigned the new value).
As r-values cannot change their value, they can be seen as the value itself (therefore the \texttt{XComputationEvent} is modelled as an local r-value memory unit, more detailed discussion goes further in current Section).

Each memory unit has an \texttt{XType} associated with it.
The X-type is a symbolic representation of the C primitive type%
\footnote{Here we should note the \porthos[2] can eventually evolve to be able to analyse programs written in an OOP language (for instance, in C++). In this case, the \texttt{XType} will have more complex structure than a simple enumeration, which it has when we need to emulate only primitive types of C language. See more detailed discussion on input language type system in Section~\ref{ch:impl:proc:x-pre-compiler:type}.} %
that is easily convertable to a SMT type (modelled as \texttt{ZType}).%TODO: add the class ZType

\paragraph{Events}
\label{ch:impl:model:xgraph:evt}

atomic action
XEventInfo

%comp event emulated by both -- optimisation?
Note, the computation-event \texttt{XComputationEvent} implements both \texttt{XEvent} and \texttt{XMemoryUnit}.
This is a model-level optimisation possible because a computation-event
%\textit{is} an event (thus it can form a node in \xgraph{}) and it also 
performs a computation over local-only memory and does not change the value of any memory unit, therefore the \textit{computation} abstraction (the CPU time) can be safely removed from the model, and the computation event can be seen as an atomic operation.

@@

DF edge: co, rf

Kinds of events: (paragraph) <pciture>

@@

Valid graph: (paragraph)
- single source, two sinks
- connected
- each node has 1 or 2 children
- branching node (2 children) instanceof XComputationEvent (evaluation of the value, no shared-memory operations)
-

@@

- numbers -- same as in gears in Figure~\ref{fig:arch}.

@@

%\textit{Type analysis} is the process of \textit{converting} the text-based type labels (obtained at variable declarations) to the type instances (modelled by the class \texttt{mousquetaires.languages.common.Type}) and \textit{propagating} them to all expressions.
%TODO: remake this paragraph considering that we have already said sth about XType
The C language has a static (resolved at compile-time) manifest (all types are declared explicitly) type system.
Comparing to languages that use type inference, the type analysis of a C program constitutes a simple propagating the type information (obtained from variables declarations) to all expressions.
%In \porthos[2], the type is modelled by the class \texttt{XType}.%TODO: + fix class and package name!
Being carried at Y and X representation levels, the type is converted to a Z-type at the Z-formula encoding stage (see Section~\ref{ch:impl:proc:z-encoder}).

%TODO: describe the Type
%TODO: maybe rename back to XType
%TODO: but definitely get rid of bit32
